\documentclass[a4paper,10pt,landscape,twocolumn]{scrartcl}

%% Settings
\newcommand\problemset{1}
\newcommand\deadline{Wednesday September 5th, 21:00h}
\newif\ifcomments
\commentsfalse % hide comments
%\commentstrue % show comments

% Packages
\usepackage[english]{exercises}
\usepackage{wasysym,hyperref}
\hypersetup{colorlinks=true, urlcolor = blue, linkcolor = blue}

\begin{document}

\practiceproblems

{\sffamily\noindent
This week's exercises deal with sets, counting and uniform probabilities. You do not have to hand these exercises in; they are optional and for practicing only. If you have questions about them, please post them to the \href{https://www.moodle.ch/lms/mod/forum/view.php?id=1634}{discussion forum} and try to help each other. We will also keep an eye on that.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{exercise}[]
	There are 11 students in a class: 4 boys and 7 girls. We need to form a group of 5 people.
	
	\begin{subex}
		How many different groups can you make?	
	\end{subex}

	 \begin{subex}
	 		How many different groups are possible with at least 4 girls?
	 \end{subex}
	 
	 \begin{subex}
		If you pick the group (uniformly) at random, what is the probability that there are at least 3 boys in the group?
	 \end{subex}
\end{exercise}

\begin{exercise}[Words]
	\begin{subex}
	How many `words' of length 5 can you make using each letter of the alphabet at most once?
	\end{subex}
	
	\begin{subex}
	And how many if the order of the letters is irrelevant? (I.e., if we treat `words'  and `sword' as the same word.)
	\end{subex}
	
	\begin{subex}
	And how many words can you make if you can use every letter as many times as you like?	
	\end{subex}

	\begin{subex}
	In how many unique ways can the letters in the word `error' be arranged?
	\end{subex}

	\begin{subex}
	 Consider a word of $n$ letters in which two letters occur more than once: $p$ and $q$ times respectively. How many unique `words' of the same length can you make of the $n$ letters?
	\end{subex}
\end{exercise}

\newpage
\begin{exercise}[Books]
	 You have 3 books on complexity theory, 2 on probability theory, and 1 novel.
	\begin{subex}
	In how many ways can the books be arranged?	
	\end{subex}
	
	\begin{subex}
	And what if	the books on complexity theory must be together but the other books can be arranged in any order?
	\end{subex}
\end{exercise}

\begin{exercise}[Poker hands]
	Calculate the probability of drawing each of these poker hands.
	
	\begin{subex}[Two-pair]
		Two cards have one rank, two cards have another rank, and the remaining card has a third rank. Example: two 2's, two 5's and a king.
	\end{subex}
	
	\begin{subex}[Three-of-a-kind]
		 Three cards have one rank and the remaining two cards have two other ranks. Example: three 2's, a five and a king.
	\end{subex}
\end{exercise}

\end{document}