\documentclass[11p,a4paper]{article}

\usepackage{amsmath, amssymb,enumerate,arev,xcolor}

\title{Board Questions, Basic Probability 2016, UvA}
\author{Teachers: Christian Schaffner, Philip Schulz; TA: Bas Cornelissen}
\date{Second Session, Sep 12th}

\begin{document}
\maketitle

\section{Rolling Dice}
\begin{enumerate}
\item Every member of your group rolls his/her 20-sided die.
\item Check if at least two of the die at your table show the same
  outcome.
\item Repeat the experiment a few time and estimate the probability of
  this event.
\item Define the sample space and probability function for this
  experiment.
\item Compute the probability of the event $A$ that at least two of
  the outcomes coincide. {\bf Hint:} It might be easier to compute the probability
  of the complement of $A$. What is this event?
\end{enumerate}

\section{Evil Squirrels}
Setting:
\begin{itemize}
\item 1.000.000 squirrels
\item 100 of them are evil, the rest is nice.
\item The proposed alarm goes of when presented with an evil squirrel 99$\%$ of the time
\item It also goes of 1$\%$ of the time when presented with a nice squirrel
\end{itemize}

Questions:
\begin{enumerate}[a)]
\item If a squirrel sets of the alarm, what's the probability that it is evil?
\item Based on this, should the evil squirrel detector be acquired?
\end{enumerate}

\section{Monty Hall}
Setting:
\begin{itemize}
\item There are 3 doors with a car behind one door and a goat behind each of the other two
\item After the candidate has chosen, one of the doors with a goat is opened
\item The candidate is given the choice to switch doors or to stick with his original choice
\end{itemize}
\paragraph{Question} Based on probabilistic calculations, should the
candidate switch doors or stick with his initial choice after one of
the goats has been revealed?
{\bf Hint:} Start off with drawing a tree and computing the
probability of winning the car when always switching the doors.
\end{document}