\documentclass[10pt, a5paper]{scrartcl}
\newcommand\problemset{5}
\newif\ifcomments
\commentsfalse % hide comments
%\commentstrue % show comments
\usepackage[english]{../../exercises}
\usepackage{amsmath,amssymb,nicefrac}
\usepackage[margin=1.7cm]{geometry}
\begin{document}
\boardquestions

\begin{exercise}[Sufficient Statistics]
  You are given a data set $ x=x_{1}^{n} $ of $ n $ independent, geometrically
  distributed observations, so $p(x_i) = \theta(1-\theta)^{x_i} $. Show that $ \sum_{i=1}^{n} x_i $ is a sufficient
  statistic for the geometric distribution.
\end{exercise}

%---------------

\begin{exercise}[Covariance]
  A coin is taken from a box containing three coins, which give heads with
  probability $p = \nicefrac{1}{3}, \nicefrac{1}{2}$, and $\nicefrac{2}{3}$.
  The mysterious coin is tossed 80 times, resulting in 49 heads and 31 tails.

  % a)
  \begin{subex}
    What is the likelihood of this data for each type of coin and which coin
    gives the maximum likelihood?
  \end{subex}
  
  % b)
  \begin{subex}
    Now suppose that we have a single coin with unknown probability $p$ of
    landing heads. Find the likelihood and log likelihood functions given the
    same data. What is the maximum likelihood estimate for $p$?
  \end{subex}
\end{exercise}

%---------------

\begin{exercise}[Dice]
  There are five fair dice each with a different number of sides: 4,6,8,12,20.
  Jon picks one of them uniformly at random rolls it and reports a 13.
  
  \begin{subex}	
    Compute the posterior probability for each die to have generated this
    outcome.
  \end{subex}
  
  \begin{subex}	
    Compute the posterior probabilities if the result had been a 5 instead. \\
    \emph{Hint: Drawing a table may help here. And please do use a calculator!}
  \end{subex}
\end{exercise}

%---------------

\begin{exercise}[Geometric maximum likelihood estimator]
  Recall that if $Y$ is  geometrically distributed, then 
  $P(Y=y) = (1-\theta)^{y}\theta$.
  You are given a set of independent, geometrically distributed observations 
  $x = x_{1}^{n}$. Find the maximum likelihood estimator of $\theta$.
\end{exercise}

\vfill\creditsboardquestions
\end{document}